---
authors: "T1 - Phoenix"
title: "US Accidents Traffic"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Libraries Used, warning=FALSE}
library(tidyverse)
library(dplyr)
library(corrplot)
library(mapview)
library(tigris)
library(sf)
library(ggplot2)
library(lubridate)
library(ezids)
```

Let us pull the dataset from the file `US_Accidents_Dec21_updated.csv` into a dataframe `df_us_acc`. As this file is huge in size, we cannot upload it on GitHub. Thus, we are storing the file in a folder called `us_accidents_dataset` that is located in one folder above the project folder.

```{r intro}
df_us_acc <- data.frame(read.csv('../us_accidents_dataset/US_Accidents_Dec21_updated.csv'))
str(df_us_acc)
```

First, let's check the percentage of NA's present in each columns of the dataset.

```{r}
(colMeans(is.na(df_us_acc)))*100
```

Here, the highest number of NAs is present in the column `Number`, followed by `Precipitation.in.`, `Wind_Chill.F.` and some other columns. As we don't require the column `Number`, we will drop the column. We have decided to keep the rest of the columns as they are part of our analysis. We are also dropping `Description` column for faster code execution.

```{r}
df_us_acc <- subset(df_us_acc, select = -c(Number, Description))
```

As we have low number of NA data for other columns, we can just drop those records.

```{r Drop NA, Factorize severity}
df_us_acc <- drop_na(df_us_acc)
```

Next, we will extract the year out of the `Start_Time` column to check the data distribution over the year.

```{r}
df_us_acc$year<-format(as.Date(df_us_acc$Start_Time, format="%Y-%m-%d"),"%Y")
ggplot(df_us_acc, aes(x = year, fill=year)) +
    geom_bar()
```

As we can see in the yearly distribution graph, the dataset has been updated with multiple data sources. Thus, we decided that the year 2021 will be the optimal subset of the data.

```{r}
clean_acc21 <- subset(df_us_acc, year==2021)
```

Let's extract the month from the `Start_Time` and check the monthly distribution.

```{r}
clean_acc21$Month<-as.numeric(format(as.Date(clean_acc21$Start_Time, format="%Y-%m-%d"),"%m"))
clean_acc21$Hour<-hour(clean_acc21$Start_Time)
```

Now, we will check the Severity distribution in the data.

```{r}
ggplot(df_us_acc, aes(x = Severity, fill=Severity)) +
    geom_bar()
```

As we have seen in the graph, the severity levels are imbalanced. We don't have a higher number of severe impacts on the traffic due to accidents as compared to the less severe. This is also true in regards to the real world. Thus, we have decided to merge level 1 & 2 into "Not Severe" & 3 & 4 into "Severe" to make our analysis more specific.

```{r}
clean_acc21 <- clean_acc21 %>% 
  mutate(Is_severe = if_else(Severity == 1 | Severity ==2 , "Not Severe", "Severe"))
clean_acc21$Is_severe <- as.factor(clean_acc21$Is_severe) # make Is_severe as factor
```

For some initial EDA, we were curious to see how the data looks on a map, particularly the DC area as we currently live here. Thus, the map below shows the accidents that took place in 2021 in the DC area.

```{r}
df_map<-dplyr::select(clean_acc21, State, Start_Lat, Start_Lng)
df_map_DC <- df_map %>% filter(State == "DC")
df_map_DC_sf <- st_as_sf(df_map_DC, coords = c("Start_Lng", "Start_Lat"), crs = 4326)
mapview(df_map_DC_sf, map.types = "Stamen.Toner",col.regions=("red"))
```




```{r Histograms for Weather Conditions}

tempHist <- ggplot(clean_acc21, aes(x=Temperature.F.)) + geom_histogram(color="black", fill = "red")+
  ggtitle("Histogram of Temperature(F) for accidents")
tempHist

windcHist <- ggplot(clean_acc21, aes(x=Wind_Chill.F.)) + geom_histogram(color="black", fill = "orange")+
  ggtitle("Histogram of Wind chill for accidents")
windcHist

humidHist <- ggplot(clean_acc21, aes(x=Humidity...)) + geom_histogram(color="black", fill = "yellow")+
  ggtitle("Histogram of Humidity for accidents")
humidHist

windsHist <- ggplot(clean_acc21, aes(x=Wind_Speed.mph.)) + geom_histogram(color="black", fill = "navy")+
  ggtitle("Histogram of Wind Speed for accidents")
windsHist

pressHist <- ggplot(clean_acc21, aes(x=Pressure.in.)) + geom_histogram(color="black", fill = "green")+
  ggtitle("Histogram of Pressure for accidents")
pressHist

visibHist <- ggplot(clean_acc21, aes(x=Visibility.mi.)) + geom_histogram(color="black", fill = "blue")+
  ggtitle("Histogram of Visibility for accidents")
visibHist

precipHist <- ggplot(clean_acc21, aes(x=Precipitation.in.)) + geom_histogram(color="black", fill = "purple")+
  ggtitle("Histogram of Precipitation for accidents")
precipHist

```

To answer the first SMART question we have, which is "Does weather affect the severity of traffic?", we wanted to check the distribution of data for numerical weather variables first.
We created histograms between numbers of accidents and weather condition elements.
From those histograms, we found that Temperature, Wind chill, and Humidity have left-skewed distributions.  
For the rest of element, which are Wind Speed, Pressure, Visibility, and Precipitation, they have quite close mean and median with a few outliers.  

```{r}
wooutlier_winds <- outlierKD2(clean_acc21, Wind_Speed.mph., rm=TRUE, boxplt=TRUE, histogram=TRUE, qqplt=TRUE)
clean_acc21_woo <- outlierKD2(wooutlier_winds, Pressure.in., rm=TRUE, boxplt=TRUE, histogram=TRUE, qqplt=TRUE)
```
So we tried to remove the outliers from Wind Speed and Pressure and without outliers, and from the generated plots we can see they are more normally distributed than the original data.
But we decided to keep the outliers because it is natural to have outliers in the weather variables as the data covers a whole year. Also the outliers does not affect the result of T-test.

```{r boxplots between Severity and wind conditions elements }

ggplot(clean_acc21, aes(x = Is_severe, y=Temperature.F.)) + 
  geom_boxplot() +
  labs(title="Temperature by Severity", x="Severity", y = "Temperature(F)")

ggplot(clean_acc21, aes(x = Is_severe, y=Wind_Chill.F.)) + 
  geom_boxplot() +
  labs(title="Wind Chill by Severity", x="Severity", y = "Wind Chill")

ggplot(clean_acc21, aes(x = Is_severe, y=Wind_Speed.mph.)) + 
  geom_boxplot() +
  labs(title="Wind Speed by Severity", x="Severity", y = "Wind Speed")

ggplot(clean_acc21, aes(x = Is_severe, y=Humidity...)) + 
  geom_boxplot() +
  labs(title="Humidity by Severity", x="Severity", y = "Humidity")

ggplot(clean_acc21, aes(x = Is_severe, y=Pressure.in.)) + 
  geom_boxplot() +
  labs(title="Pressure by Severity", x="Severity", y = "Pressure")

ggplot(clean_acc21, aes(x = Is_severe, y=Visibility.mi.)) + 
  geom_boxplot() +
  labs(title="Visibility by Severity", x="Severity", y = "Visibility")

ggplot(clean_acc21, aes(x = Is_severe, y=Precipitation.in.)) + 
  geom_boxplot() +
  labs(title="Precipitation by Severity", x="Severity", y = "Precipitation")

## The same variables(Temp, wind chill, humidity) have pretty plots
```
We tried to see the distribution of weather elements by two different severity levels which are 'Severe' and 'Not Severe'. 

For Temperature, Wind chill, and Humanity, we can see the difference on range of data distribution and outliers by severity levels.

For the rest of element such as wind speed, pressure, visibility and precipitation, they still do not have a wide range of data but we can see the distribution by two severity levels more conveniently with boxplots.
```{r T-test on Severity and weather condition elements}
box_clean2_severe = subset(clean_acc21, Is_severe == 'Severe')
box_clean2_notsevere = subset(clean_acc21, Is_severe == 'Not Severe')
print("Temperature.F")
t.test(box_clean2_severe$Temperature.F., box_clean2_notsevere$Temperature.F.)
print("Wind_Chill.F.")
t.test(box_clean2_severe$Wind_Chill.F., box_clean2_notsevere$Wind_Chill.F.)
print("Humidity...")
t.test(box_clean2_severe$Humidity..., box_clean2_notsevere$Humidity...)
print("Wind_Speed.mph.")
t.test(box_clean2_severe$Wind_Speed.mph., box_clean2_notsevere$Wind_Speed.mph.)
print("Visibility.mi.")
t.test(box_clean2_severe$Visibility.mi., box_clean2_notsevere$Visibility.mi.)
print("Pressure.in.")
t.test(box_clean2_severe$Pressure.in., box_clean2_notsevere$Pressure.in.)
print("Precipitation.in.")
t.test(box_clean2_severe$Precipitation.in., box_clean2_notsevere$Precipitation.in.)
```
First, I divided the data into two different data by subsetting by the severity to check the means of Weather elements between two different severity levels will be same or not. 
Then we performed the two-sample t-test on Severity and weather elements since the weather numerical variables are quantitative and we have two samples based on the severity levels. 

H0: The means of Temperature/WindChill/Humidity/Wind Speed/Pressure/Visibility/Precipitation will be same between different Severity levels.
H1: The means of Temperature/WindChill/Humidity/Wind Speed/Pressure/Visibility/Precipitation will NOT be same between different Severity levels.

The p-value from all tests except for Visibility are lower than 0.05 so we can reject the H0 for every weather variables but Visibility, which means that means from weather variables except for Visibility were different by its severity level of traffic. 

From these t-tests, we can conclude that numerical weather variables such as Temperature, WindChill, Humidity, Pressure, Wind Speed and Precipitation affect the severity of traffic.

However, numerical weather variables are not only variables from weather conditions.
There are categorical weather variables in our dataset, such as Wind Directions and Weather Conditions. 
```{r Wind_Direction}
# wind direction with severity
ggplot(clean_acc21, aes(Wind_Direction, ..prop.., group = Is_severe)) +
  geom_bar(aes(fill = Is_severe)) +
  #scale_y_continuous(labels = percent) +
  labs(x = "Wind Direction",
       y = "Proportion",
       title = "Wind direction by Severity") +
  theme(text = element_text(size=8)) 
```
We made a bar plot to see the distribution of the Severity by wind directions.
As the bar plot between Severity and Wind direction here shows the similar distribution for each severity levels on most of wind direction. So we can infer that wind direction does not affect much on the severity of traffic and decided not to perform any statistical analysis on the Wind Direction variable.
```{r Weather_Condition variables}
# Weather Condition percentage
unique(clean_acc21$Weather_Condition)
WC <- clean_acc21 %>%
  group_by(Weather_Condition) %>%
  summarise(cnt = n()) %>%
  mutate(freq = (round(cnt/sum(cnt), 3))*100 )%>%
  arrange(desc(freq)) %>%
  filter(freq > 1)
WC
WC %>%
    ggplot() +
    geom_col(mapping = aes(x=reorder(Weather_Condition, -freq), y=freq, fill = Weather_Condition)) +
    labs(x = "Weather Condition", y="%", title ="Top 8 Weather Conditions with accidents") +
    theme(text = element_text(size=7))
```
Secondly, we took a look at the Weather Condition variable.  
There are so many weather conditions in our dataset, so I tried to make a barplot with the top 8 weather conditions when accidents happened. The major weather conditions when car accidents happened were 'Fair', 'Cloudy', 'Mostly Cloudy', 'Partly Cloudy', 'Light Rain', 'Fog', 'Light Snow', and 'Haze'.  

Now we can see that the most frequent weather was 'Fair', but since the Weather Conditions variable is divided into detailed conditions as you can see here with cloudy, mostly cloudy, and partly cloudy, so we decided to take a chi-squared test on all weather conditions and severity. 

```{r}
# Try the Chi-Squared Test on all Weather_Conditions and Severity
WCtable <- table(clean_acc21$Weather_Condition , clean_acc21$Severity)
xkabledply(WCtable, title = "Severity by Weather Conditions")
chitest = chisq.test(WCtable)
chitest
```
To identify if the severity of traffic dependent on Weather Condition, we took a Chi-Squared test. As Weather Condition is a categorical variables, and we treated the severity in an original form which is numerical so we can see the dependency between weather conditions and the severity using a chi-squared test.  

H0 : Severity and weather conditions are independent.
H1 : Severity and weather conditions are NOT independent.  
As we can see here, the the P-value from Chi-squared test is lower than 0.05 for Weather Conditions variable so we can reject the H0. Which means the Severity of traffic and weather conditions are Dependent. 

From these analysis on numerical and categorical weather variables, we can answer our SMART question about the impact for weather on the severity of traffic.
We concluded that the numerical weather condition elements except for visibility affect the severity of traffic.
However, the wind direction does not affect much the severity of traffic because it does not show differences on severity by each of directions. 
For the weather conditions variables, we can observe that weather conditions at the time accidents happened affect the severity of traffic.  


SMART QUESTION: 

Do Nearby Road Elements affect the severity of traffic?

To determine whether nearby road components have an impact on the severity of the traffic, we have conducted the exploratory data analysis listed below.

```{r}
clean_acc21 %>% 
      group_by(Amenity, Bump, Crossing, Give_Way, Junction, No_Exit, Railway, Roundabout, Station, Stop, Traffic_Calming, Traffic_Signal, Turning_Loop)%>%
    summarise(percentage = n()/nrow(clean_acc21) *100 )%>%
    arrange(-percentage) %>%
    filter(percentage > 1) -> accidents_per_roadelement

head(accidents_per_roadelement)
```
Here, we've compiled a list of incidents where surrounding road features like an amenity, a bump, a crossing, a give-way, a junction, a no-exit, a railroad, a roundabout, a station, a stop, a traffic signal, or a turning loop contributed to the accident.

```{r}


acc_road_element_per <- tibble(c("None", "Junction","Crossing", "Traffic signal", "Crossing and traffic signal", "Station", "Stop" ), pull(accidents_per_roadelement, percentage), .name_repair = ~ c("road_elements", "percentage"))
acc_road_element_per
       
```

The tibble above illustrates accidents that happened as a result of local road elements in areas where there are more than 1% of road accidents.

```{r fig.align="center", echo = FALSE,fig.width = 12}
options(repr.plot.width = 30, repr.plot.height = 8)
acc_road_element_per %>%
    ggplot() +
    geom_col(mapping = aes(x=reorder(road_elements, -percentage), y=percentage, fill = road_elements)) +
    
  labs(title="Accidents By Nearby road elements", x="Road Elements", y = "Percentage")+
  theme(plot.title = element_text(hjust = 0.5))

```
The graph above shows the accidents that happened as a result of surrounding road features.

```{r}


accidents_per_roadelement<-clean_acc21 %>% 
      select(Is_severe,Junction,Crossing,Stop,Station,Traffic_Signal)


head(accidents_per_roadelement)

```

Here, we've created a data frame with columns for junction, crossing, stop, station, and traffic signal, as well as a severity rating.

```{r}
Junction_data<- subset(accidents_per_roadelement, Junction=="True",
select=c(Is_severe,Junction))
head(Junction_data)

Crossing_data<- subset(accidents_per_roadelement, Crossing=="True",
select=c(Is_severe,Crossing))
head(Crossing_data)

Stop_data<- subset(accidents_per_roadelement, Stop=="True",
select=c(Is_severe,Stop))
head(Stop_data)

Station_data<- subset(accidents_per_roadelement, Station=="True",
select=c(Is_severe,Station))
head(Station_data)

Traffic_Signal_data<- subset(accidents_per_roadelement, Traffic_Signal=="True",
select=c(Is_severe,Traffic_Signal))
head(Traffic_Signal_data)
```

Here, the data frame is divided based on the accident's severity level and the surrounding road element that caused it.

```{r}
ggplot(data=Junction_data, aes(x=Is_severe, y=Junction, fill=Is_severe)) +
  geom_bar(stat="identity")+
  labs(title="Accidents By Junction", x="Severity", y = "Junction")+
  theme(plot.title = element_text(hjust = 0.5))

ggplot(data=Crossing_data, aes(x=Is_severe, y=Crossing, fill=Is_severe)) +
  geom_bar(stat="identity")+
  labs(title="Accidents By Crossing", x="Severity", y = "Crossing")+
  theme(plot.title = element_text(hjust = 0.5))

ggplot(data=Stop_data, aes(x=Is_severe, y=Stop, fill=Is_severe)) +
  geom_bar(stat="identity")+
  labs(title="Accidents By Stop", x="Severity", y = "Stop")+
  theme(plot.title = element_text(hjust = 0.5))

ggplot(data=Station_data, aes(x=Is_severe, y=Station, fill=Is_severe)) +
  geom_bar(stat="identity")+
  labs(title="Accidents By Station", x="Severity", y = "Station")+
  theme(plot.title = element_text(hjust = 0.5))

ggplot(data=Traffic_Signal_data, aes(x=Is_severe, y=Traffic_Signal, fill=Is_severe)) +
  geom_bar(stat="identity")+
  labs(title="Accidents By Traffic Signal", x="Severity", y = "Traffic Sisgnal")+
  theme(plot.title = element_text(hjust = 0.5))
```

The aforementioned graphs show the degree of severity caused by neighboring road elements such junctions, crossings, stops, stations, and traffic signals.

```{r}

accidents_per_roadelement <- accidents_per_roadelement %>% 
  mutate(severe_num = if_else(Is_severe== "Severe", 1, 0))
accidents_per_roadelement <- accidents_per_roadelement %>% 
  mutate(Crossing_num = if_else(Crossing== "True", 1, 0))
accidents_per_roadelement <- accidents_per_roadelement %>% 
  mutate(Stop_num = if_else(Stop== "True", 1, 0))
accidents_per_roadelement <- accidents_per_roadelement %>% 
  mutate(Station_num = if_else(Station== "True", 1, 0))
accidents_per_roadelement <- accidents_per_roadelement %>% 
  mutate(Traffic_Signal_num = if_else(Traffic_Signal== "True", 1, 0))
accidents_per_roadelement <- accidents_per_roadelement %>% 
  mutate(Junction_num = if_else(Junction== "True", 1, 0))

```

In order to run the anova test, we are turning the category data presented here into numerical data.

```{r}
Junction_anova = aov(Junction_num~severe_num, data=accidents_per_roadelement)
summary(Junction_anova)

Crossing_anova = aov(Crossing_num~severe_num, data=accidents_per_roadelement)
summary(Crossing_anova)

Stop_anova = aov(Stop_num~severe_num, data=accidents_per_roadelement)
summary(Stop_anova)

Station_anova = aov(Station_num~severe_num, data=accidents_per_roadelement)
summary(Station_anova)

Traffic_signal_anova = aov(Traffic_Signal_num~severe_num, data=accidents_per_roadelement)
summary(Traffic_signal_anova)
```

Since the outcome variable is quantitative and the predictor variable is categorical, we are doing an anova test.
H0 : The Severity of traffic and Nearby Road Elements are independent
H1 : The Severity of traffic and  Nearby Road Elements are NOT independent
The P-value from Anova test is lower than 0.05 for all the Nearby Road Elements variable so we can reject the H0.
→ Nearby Road Elements Junction,Crossing, Stop, Station and Traffic Signal affects the Severity of Traffic.



########### ADD SOME COMMENTS HERE TO START THE 3RD QUESTION


```{r}
ggplot(clean_acc21, aes(x = Month, fill=Month,)) +geom_bar()+labs(title="Accident Distribution over the Months in 2021", x="Months", y = "count")+
  theme(plot.title = element_text(hjust = 0.5))
```
The Graph above shows frequency of accidents in 2021 during different Months.
```{r}
ggplot(clean_acc21, aes(x = Hour, fill=as.factor(Hour))) + geom_bar()+
  labs(title="Accident Distribution in Hours ", x="Hours", y = "count")+
  theme(plot.title = element_text(hjust = 0.5))
```
The Graph above shows the frequency of accidents in 2021 during different hours of the day.

```{r}
hist(clean_acc21$Hour,main="Histogram for Hours",xlab = "Hours")
```

```{r}
test_Hour <- chisq.test(table(clean_acc21$Severity, clean_acc21$Hour))
test_Hour
```
Took a Chi-Squared test to see if the severity and Hour of day are independent.
We have performed Chi-Squared test because both variables are categorical.
H0 : Severity and Hour are independent
H1 : Severity and Hour are NOT independent
We can reject the H0. 
```{r}
test_Month <- chisq.test(table(clean_acc21$Severity, clean_acc21$Month))
test_Month
```
Took a Chi-Squared test to see the severity and weather Months are independent.
We have performed Chi-Squared test because both variables are categorical.
H0 : Severity and Month are independent
H1 : Severity and Month are NOT independent
We can reject the H0. 


```{r}
cor(clean_acc21$Severity, clean_acc21$Month)
```
Correlation test performed to test strength of relationship between severity and Month.

```{r}
cor(clean_acc21$Severity, clean_acc21$Hour)
```
Correlation test performed to test strength of relationship between severity and Hour.

