---
authors: 
title: "us_accidents_traffic"
output: html_document
date: "2022-10-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Libraries Used}
library(tidyverse)
library(dplyr)
library(corrplot)
library(mapview)
library(tigris)
library(sf)
library(ggplot2)
library(lubridate)
library(ezids)
```

```{r intro}
df_us_acc <- data.frame(read.csv('../us_accidents_dataset/US_Accidents_Dec21_updated.csv'))
str(df_us_acc)
```

```{r}
(colMeans(is.na(df_us_acc)))*100
```


```{r}
df_us_acc$year<-format(as.Date(df_us_acc$Start_Time, format="%Y-%m-%d"),"%Y")
head(df_us_acc)
```

```{r}
ggplot(df_us_acc, aes(x = year, fill=year)) +
    geom_bar()
```


```{r}
accidents_2021 <- subset(df_us_acc, year==2021)
head(accidents_2021)
```

```{r}
accidents_2021$month<-format(as.Date(accidents_2021$Start_Time, format="%Y-%m-%d"),"%m")
head(accidents_2021)
```

```{r}
accidents_2021<-subset(accidents_2021, select = -c(Number, ID, Description))
```

```{r}
(colMeans(is.na(accidents_2021)))*100
```

```{r}
accidents_2021 <- accidents_2021 %>% 
  mutate(Is_severe = if_else(Severity == 1 | Severity ==2 , "Not Severe", "Severe"))
```

```{r}
accidents_2021  %>%
  summarise_all(list(~is.na(.)))%>%
  pivot_longer(everything(),
               names_to = "variables", values_to="missing") %>%
  count(variables, missing) %>%
  ggplot(aes(y=variables,x=n,fill=missing))+
  geom_col()
```



```{r Drop NA, Factorize severity}
clean_acc21 <- drop_na(accidents_2021)

```

```{r}
clean_acc21$Is_severe <- as.factor(clean_acc21$Is_severe) # make Is_severe as factor
str(clean_acc21)
```

```{r}
clean_acc21  %>%
  summarise_all(list(~is.na(.)))%>%
  pivot_longer(everything(),
               names_to = "variables", values_to="missing") %>%
  count(variables, missing) %>%
  ggplot(aes(y=variables,x=n,fill=missing))+
  geom_col()
```

```{r}
us_date<-clean_acc21$Start_Time 
us_date_formatted<- as.Date(us_date,format="%Y-%m-%d")
clean_acc21$Month<-as.numeric(format(us_date_formatted, "%m"))
```

```{r}
us_time<- clean_acc21$Start_Time
clean_acc21$Hour<-hour(us_time)
```

```{r}
ggplot(clean_acc21, aes(x = month, fill=month,)) +geom_bar()+labs(title="Accident Distribution over the Months in 2021", x="Months", y = "count")+
  theme(plot.title = element_text(hjust = 0.5))
ggplot(clean_acc21, aes(x = Hour, fill=as.factor(Hour))) + geom_bar()+
  labs(title="Accident Distribution in Hours ", x="Hours", y = "count")+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
hist(clean_acc21$Hour,main="Histogram for Hours",xlab = "Hours")
```

```{r}
df_map<-dplyr::select(clean_acc21, State, Start_Lat, Start_Lng)
head(df_map)
```

```{r}
df_map_DC <- df_map %>% filter(State == "DC")

df_map_DC %>% glimpse()
```

```{r}
df_map_DC_sf <- st_as_sf(df_map_DC, coords = c("Start_Lng", "Start_Lat"), crs = 4326)
```

```{r}
mapview(df_map_DC_sf, main="Map for Accidents in DC(2021)" ,xcol = "Start_Lng", ycol = "Start_Lat", grid = FALSE,col.regions=("red"))
```

```{r}
mapview(df_map_DC_sf, map.types = "Stamen.Toner",col.regions=("red"))
```

```{r Correlations for numerical values}
# subsetting numerical columns only to check the corelationship between numerical variables
loadPkg("corrplot")
num_clean_acc21 <- select_if(clean_acc21, is.numeric)
str(num_clean_acc21)
cor_num_clean_acc21 <- cor(num_clean_acc21)
cor_num_clean_acc21
corrplot(cor_num_clean_acc21, method = "number")
```

```{r Histograms for Weather Conditions}

tempHist <- ggplot(clean_acc21, aes(x=Temperature.F.)) + geom_histogram(color="black", fill = "red")+
  ggtitle("Histogram of Temperature(F) for accidents")
tempHist

windcHist <- ggplot(clean_acc21, aes(x=Wind_Chill.F.)) + geom_histogram(color="black", fill = "orange")+
  ggtitle("Histogram of Wind chill for accidents")
windcHist

humidHist <- ggplot(clean_acc21, aes(x=Humidity...)) + geom_histogram(color="black", fill = "yellow")+
  ggtitle("Histogram of Humidity for accidents")
humidHist

windsHist <- ggplot(clean_acc21, aes(x=Wind_Speed.mph.)) + geom_histogram(color="black", fill = "navy")+
  ggtitle("Histogram of Wind Speed for accidents")
windsHist

pressHist <- ggplot(clean_acc21, aes(x=Pressure.in.)) + geom_histogram(color="black", fill = "green")+
  ggtitle("Histogram of Pressure for accidents")
pressHist

visibHist <- ggplot(clean_acc21, aes(x=Visibility.mi.)) + geom_histogram(color="black", fill = "blue")+
  ggtitle("Histogram of Visibility for accidents")
visibHist

precipHist <- ggplot(clean_acc21, aes(x=Precipitation.in.)) + geom_histogram(color="black", fill = "purple")+
  ggtitle("Histogram of Precipitation for accidents")
precipHist

```

First, we looked at the histogram between numbers of accidents and weather condition elements to check distribution of data.
For Temperature, Wind chill, and Humanity, each of distributions is left-skewed. 
For the rest of element, they have quite close mean and median and a few outliers. But we decided to keep the outliers because it is natural to have outliers in the weather condition elements as it covers a whole year. Also the outliers does not affect the result of T-test. 

```{r boxplots between Severity and wind conditions elements }

ggplot(clean_acc21, aes(x = Is_severe, y=Temperature.F.)) + 
  geom_boxplot() +
  labs(title="Temperature by Severity", x="Severity", y = "Temperature(F)")

ggplot(clean_acc21, aes(x = Is_severe, y=Wind_Chill.F.)) + 
  geom_boxplot() +
  labs(title="Wind Chill by Severity", x="Severity", y = "Wind Chill")

ggplot(clean_acc21, aes(x = Is_severe, y=Wind_Speed.mph.)) + 
  geom_boxplot() +
  labs(title="Wind Speed by Severity", x="Severity", y = "Wind Speed")

ggplot(clean_acc21, aes(x = Is_severe, y=Humidity...)) + 
  geom_boxplot() +
  labs(title="Humidity by Severity", x="Severity", y = "Humidity")

ggplot(clean_acc21, aes(x = Is_severe, y=Pressure.in.)) + 
  geom_boxplot() +
  labs(title="Pressure by Severity", x="Severity", y = "Pressure")

ggplot(clean_acc21, aes(x = Is_severe, y=Visibility.mi.)) + 
  geom_boxplot() +
  labs(title="Visibility by Severity", x="Severity", y = "Visibility")

ggplot(clean_acc21, aes(x = Is_severe, y=Precipitation.in.)) + 
  geom_boxplot() +
  labs(title="Precipitation by Severity", x="Severity", y = "Precipitation")

## The same variables(Temp, wind chill, humidity) have pretty plots

```

```{r T-test on Severity and weather condition elements}
box_clean2_severe = subset(clean_acc21, Is_severe == 'Severe')
box_clean2_notsevere = subset(clean_acc21, Is_severe == 'Not Severe')
print("Temperature.F")
t.test(box_clean2_severe$Temperature.F., box_clean2_notsevere$Temperature.F.)
print("Wind_Chill.F.")
t.test(box_clean2_severe$Wind_Chill.F., box_clean2_notsevere$Wind_Chill.F.)
print("Humidity...")
t.test(box_clean2_severe$Humidity..., box_clean2_notsevere$Humidity...)
print("Wind_Speed.mph.")
t.test(box_clean2_severe$Wind_Speed.mph., box_clean2_notsevere$Wind_Speed.mph.)
print("Visibility.mi.")
t.test(box_clean2_severe$Visibility.mi., box_clean2_notsevere$Visibility.mi.)
print("Pressure.in.")
t.test(box_clean2_severe$Pressure.in., box_clean2_notsevere$Pressure.in.)
print("Precipitation.in.")
t.test(box_clean2_severe$Precipitation.in., box_clean2_notsevere$Precipitation.in.)
```
H0: The means of Temperature/WindChill/Humidity/Wind Speed/Pressure/Visibility/Precipitation will be same between different Severity levels.
H1: The means of Temperature/WindChill/Humidity/Wind Speed/Pressure/Visibility/Precipitation will NOT be same between different Severity levels.
The P-value for the all are lower than 0.05 for those all so we can reject the H0 except for Precipitation.
```{r Wind_Direction}
# wind direction with severity
ggplot(clean_acc21, aes(Wind_Direction, ..prop.., group = Is_severe)) +
  geom_bar(aes(fill = Is_severe)) +
  #scale_y_continuous(labels = percent) +
  labs(x = "Wind Direction",
       y = "Proportion",
       title = "Wind direction by Severity") +
  theme(text = element_text(size=8)) 
```
As the bar plot between Severity and Wind direction here shows the similar distribution for each severity levels on most of wind direction, we can conclude that wind direction does not affect much on the severity of traffic.
```{r Weather_Condition variables}
# Weather Condition percentage
unique(clean_acc21$Weather_Condition)
WC <- clean_acc21 %>%
  group_by(Weather_Condition) %>%
  summarise(cnt = n()) %>%
  mutate(freq = (round(cnt/sum(cnt), 3))*100 )%>%
  arrange(desc(freq)) %>%
  filter(freq > 1)
WC
WC %>%
    ggplot() +
    geom_col(mapping = aes(x=reorder(Weather_Condition, -freq), y=freq, fill = Weather_Condition)) +
    labs(x = "Weather Condition", y="%", title ="Top 8 Weather Conditions with accidents") +
    theme(text = element_text(size=7))
```

```{r}
# Try the Chi-Squared Test on all Weather_Conditions and Severity
WCtable <- table(clean_acc21$Weather_Condition , clean_acc21$Severity)
xkabledply(WCtable, title = "Severity by Weather Conditions")
chitest = chisq.test(WCtable)
chitest
```
Took a Chi-Squared test to see the severity and weather conditions are independent.
H0 : Severity and weather conditions are independent
H1 : Severity and weather conditions are NOT independent
We can reject the H0. 


```{r}
clean_acc21 %>% 
      group_by(Amenity, Bump, Crossing, Give_Way, Junction, No_Exit, Railway, Roundabout, Station, Stop, Traffic_Calming, Traffic_Signal, Turning_Loop)%>%
    summarise(percentage = n()/nrow(clean_acc21) *100 )%>%
    arrange(-percentage) %>%
    filter(percentage > 1) -> accidents_per_roadelement

head(accidents_per_roadelement)
```



```{r}


acc_road_element_per <- tibble(c("None", "Junction","Crossing", "Traffic signal", "Crossing and traffic signal", "Station", "Stop" ), pull(accidents_per_roadelement, percentage), .name_repair = ~ c("road_elements", "percentage"))
acc_road_element_per
       
```




```{r fig.align="center", echo = FALSE,fig.width = 12}
options(repr.plot.width = 30, repr.plot.height = 8)
acc_road_element_per %>%
    ggplot() +
    geom_col(mapping = aes(x=reorder(road_elements, -percentage), y=percentage, fill = road_elements)) +
    
  labs(title="Accidents By Nearby road elements", x="Road Elements", y = "Percentage")+
  theme(plot.title = element_text(hjust = 0.5))

```

```{r}


accidents_per_roadelement<-clean_acc21 %>% 
      select(Is_severe,Junction,Crossing,Stop,Station,Traffic_Signal)


head(accidents_per_roadelement)

```

```{r}
Junction_data<- subset(accidents_per_roadelement, Junction=="True",
select=c(Is_severe,Junction))
head(Junction_data)
```
```{r}
ggplot(data=Junction_data, aes(x=Is_severe, y=Junction, fill=Is_severe)) +
  geom_bar(stat="identity")+
  labs(title="Accidents By Junction", x="Severity", y = "Junction")+
  theme(plot.title = element_text(hjust = 0.5))
```




```{r}
Crossing_data<- subset(accidents_per_roadelement, Crossing=="True",
select=c(Is_severe,Crossing))
head(Crossing_data)
```
```{r}
ggplot(data=Crossing_data, aes(x=Is_severe, y=Crossing, fill=Is_severe)) +
  geom_bar(stat="identity")+
  labs(title="Accidents By Crossing", x="Severity", y = "Crossing")+
  theme(plot.title = element_text(hjust = 0.5))
```


```{r}
Stop_data<- subset(accidents_per_roadelement, Stop=="True",
select=c(Is_severe,Stop))
head(Stop_data)
```
```{r}
ggplot(data=Stop_data, aes(x=Is_severe, y=Stop, fill=Is_severe)) +
  geom_bar(stat="identity")+
  labs(title="Accidents By Stop", x="Severity", y = "Stop")+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
Station_data<- subset(accidents_per_roadelement, Station=="True",
select=c(Is_severe,Station))
head(Station_data)
```
```{r}
ggplot(data=Station_data, aes(x=Is_severe, y=Station, fill=Is_severe)) +
  geom_bar(stat="identity")+
  labs(title="Accidents By Station", x="Severity", y = "Station")+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
Traffic_Signal_data<- subset(accidents_per_roadelement, Traffic_Signal=="True",
select=c(Is_severe,Traffic_Signal))
head(Traffic_Signal_data)
```
```{r}
ggplot(data=Traffic_Signal_data, aes(x=Is_severe, y=Traffic_Signal, fill=Is_severe)) +
  geom_bar(stat="identity")+
  labs(title="Accidents By Traffic Signal", x="Severity", y = "Traffic Sisgnal")+
  theme(plot.title = element_text(hjust = 0.5))
```



```{r}

accidents_per_roadelement <- accidents_per_roadelement %>% 
  mutate(severe_num = if_else(Is_severe== "Severe", 1, 0))
accidents_per_roadelement <- accidents_per_roadelement %>% 
  mutate(Crossing_num = if_else(Crossing== "True", 1, 0))
accidents_per_roadelement <- accidents_per_roadelement %>% 
  mutate(Stop_num = if_else(Stop== "True", 1, 0))
accidents_per_roadelement <- accidents_per_roadelement %>% 
  mutate(Station_num = if_else(Station== "True", 1, 0))
accidents_per_roadelement <- accidents_per_roadelement %>% 
  mutate(Traffic_Signal_num = if_else(Traffic_Signal== "True", 1, 0))
accidents_per_roadelement <- accidents_per_roadelement %>% 
  mutate(Junction_num = if_else(Junction== "True", 1, 0))

```

```{r}
Junction_anova = aov(Junction_num~severe_num, data=accidents_per_roadelement)
summary(Junction_anova)
```


```{r}
Crossing_anova = aov(Crossing_num~severe_num, data=accidents_per_roadelement)
summary(Crossing_anova)
```


```{r}
Stop_anova = aov(Stop_num~severe_num, data=accidents_per_roadelement)
summary(Stop_anova)
```
```{r}
Station_anova = aov(Station_num~severe_num, data=accidents_per_roadelement)
summary(Station_anova)
```
```{r}
Traffic_signal_anova = aov(Traffic_Signal_num~severe_num, data=accidents_per_roadelement)
summary(Traffic_signal_anova)
```

```{r}
test_Hour <- chisq.test(table(clean_acc21$Severity, clean_acc21$Hour))
test_Hour
```

```{r}
test_Month <- chisq.test(table(clean_acc21$Severity, clean_acc21$Month))
test_Month
```

```{r}
cor(clean_acc21$Severity, clean_acc21$Month)
```

```{r}
cor(clean_acc21$Severity, clean_acc21$Hour)
```
